---
title: "**Data visualization with R** - `coronavirus` and `ggplot2` Packages"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    code_fold: show
---

```{r setup,echo=FALSE, include=TRUE, warning=FALSE}
# =============== SETTING ENVIROMENT VARIABLES ===============

Sys.setenv("LANGUAGE"="En", TZ = "UTC")

knitr::opts_chunk$set(echo = TRUE)

klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done',
               color = 'darkred', position = "right")
```
<style>
body {
text-align: justify}
</style>

# Coronavirus

As we all know the coronavirus is a current issue that concerns all humanity because of its gravity. Wanting to know more, I found a package in CRAN that helped me understand the spread of the disease globally. Although for many it may seem like a rudimentary table of daily case records, I found in this simple data set an opportunity to do what I like most - graphics - and to share with those people who want to venture to use **R** something of what I have learned with `ggplot`. Of course, without neglecting how serious this disease is and how it has been its temporal and spatial evolution on the planet. Well, as the saying goes:

> *A picture is worth a thousand words.*

Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) has released a package at CRAN which provides a daily summary of the **Coronavirus (COVID-19)** cases by state/province. The package is available at <https://cran.r-project.org/web/packages/coronavirus/index.html>. Below, I'll show some simple analysis with this dataset with the help of *ggplot2* in *R*, my aim for next posts is to show you how to create animated plots with *gganimate* package.

The main gist of the process is to show how coronavirus is spread in the World nowadays, and how the cases in each country are increasing. To get started, you’ll need to download coronavirus from Github using the `devtools` package:

```{r, echo=T, message=F}
# *************** Install package ***************

if (!require(devtools)) {
    install.packages("devtools")
}
if (!require(coronavirus)) {
  devtools::install_github("RamiKrispin/coronavirus")
}
```

In addition to this package, I’ll also use `dplyr` and `magrittr` for data manipulation, `ggplot2` for plotting, and `maps` for manipulate geographical data. 

```{r , echo=T, message=F}
library(coronavirus)  # The 2019 Novel Coronavirus COVID-19 (2019-nCoV) Dataset
library(dplyr)        # A Grammar of Data Manipulation
library(magrittr)     # A Forward-Pipe Operator for R
library(maps)         # Draw Geographical Maps
library(ggplot2)      # Create Elegant Data Visualisations Using the Grammar of Graphics
```

### COVID-19 spread evolution around the Globe

When loading the `library(coronavirus)` we get the dataset which contains the daily summary of Coronavirus cases (confirmed, death, and recovered), by state/province for some countries as seen below:

```{r, layout="l-body-outset"}
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
DT::datatable(coronavirus, rownames = FALSE)
```

Now we are going to calculate the quantity of cases confirmed, death, and recovered until the last time dataset was updated for each country. Grouping data by Country.Region, Province.State, Lat, Long and type and adding up the cases, we have:

```{r}
coronavirus_data <- 
coronavirus %>% 
  group_by(Country.Region, Province.State, Lat, Long, type) %>% 
  summarise(total = sum(cases)) %>% 
  rename(Country = Country.Region)

coronavirus_data %>% head() %>% kable()
```

Now we are going to locate each coordinate of the dataset in a worldmap using a point for each one. Color and size of the points are going to be related with the quantity of cases, we are going to build a grid of plots with the `facet_grid` function to create a map for each type of case:

```{r fig.height=6,fig.width=9, fig.align = "center", out.width="100%", class.output="scroll-100"}

# *************** Create a Map theme ***************
theme.maptheeme <-
  theme(text = element_text(family = "Gill Sans", color = "#444444")) +
  theme(plot.title = element_text(size = 16, face = "bold")) +
  theme(plot.subtitle = element_text(size = 12)) +
  theme(panel.grid = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.title = element_blank()) +
  theme(legend.background = element_blank()) +
  theme(legend.key = element_blank()) +
  theme(legend.title = element_text(size = 12)) +
  theme(legend.text = element_text(size = 10)) +
  theme(legend.direction = "vertical") +
  theme(legend.box = "horizontal") +
  theme(panel.background = element_rect(fill = "#2B343D")) +
  theme(panel.grid = element_blank()) +
  theme(panel.border = element_rect(color = "#2B343D", fill = NA)) +
  theme(strip.background = element_rect(color = "#2B343D", fill = "#2B343D")) +
  theme(strip.text = element_text(colour = "lightgray")) +
  theme(plot.title = element_text(face = "bold"))

# *************** Convert type cases to capital letters ***************
coronavirus_data$type <- 
  factor(coronavirus_data$type, labels = toupper(unique(coronavirus_data$type))) # To upper case

# *************** Get the woldmap ***************

map.world <- map_data("world")  # Here we are obtaining the worldmap data, borders, countries, etc.

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group), 
               fill="#222524", colour = "#707070", size = 0.1, alpha = 0.5) +   # World map
  geom_point(data = coronavirus_data, 
             aes(x = Long, y = Lat, size = total, fill = type, color = type),
             shape = 21, stroke =0.4, alpha = 0.4) +                            # Coronavirus geometry
  facet_wrap(facets = ~type, ncol = 2) +                                        # Facet
  scale_size_continuous(range = c(.7,15), 
                        labels = scales::comma_format()) +              
  scale_fill_manual(values = c("#ffe146", "#be2026", "#21b6ba")) +
  scale_color_manual(values = c("#ffe146", "#be2026", "#21b6ba"), guide=FALSE) +# Scales
  labs(x="", y="", title = "COVID-19", 
       subtitle = paste("Total number of coronavirus cases from the origin of the disease to",
                        max(coronavirus$date)),
       size = "Confirmed \ncases" , fill = "Type\n") +                          # Aesthetics
  guides(fill = guide_legend(override.aes = list(size=10, alpha = 0.6)),
         size = guide_legend(override.aes = list(alpha = 0.9, stroke = 1))) +
  coord_fixed() +
  theme.maptheeme +
  theme(legend.position = c(0.75, 0.25))
```

### Countries where COVID-19 has reached

Now it is time to track the virus from China. For that I create a dataframe with the Chinese province where were registered the first cases, i.e. in *Wuhan*, capital of *Hubei* province. Then I created another dataframe representing places where COVID-19 came. Once this is done, I selected only the Lat and Long and represented only once for each country and to avoid overlapping of lines. I created a random column named curvature so that the trajectories seem more real on the final map.

```{r fig.height=5,fig.width=9, fig.align = "center", out.width="100%"}
# *************** Get the origin points to the curves ***************
# Origin
Wuhan <- 
data.frame(Province.State = "Hubei",
           Lat = 30.97564,
           Long = 112.2707)

# Destination of the virus
migr_corona <- 
coronavirus %>% filter(Province.State != "Hubei") %>% 
  select(Lat, Long) %>% distinct() %>% 
  mutate(curvature = runif(n(), min = -0.4, max = 0.4)) # I created a random curvature index

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group),
               fill="#222524", colour = "#707070", size = 0.08, alpha = 0.5) +  # World map
lapply(split(migr_corona, 1:nrow(migr_corona)), function(dat) {
    geom_curve(data = dat, aes(x=Wuhan$Long, y=Wuhan$Lat, xend=Long, yend=Lat),
               color= alpha("#21b6ba", 0.8),
               size=.1,
               curvature = dat["curvature"]) }) +                               # Conect points
  geom_point(data=Wuhan,
             aes(x=Long, y=Lat), 
             fill="red", shape = 21, stroke=0.2,
             color ="white", size=1.5) +
  geom_point(data=migr_corona,
             aes(x=Long, y=Lat), 
             color =alpha("#21b6ba", 0.9),size=0.2) +
  coord_fixed() +
  guides(color=guide_legend(nrow = 3)) +
  labs(x="", y="", title = "COVID-19", subtitle = "Countries where the virus has reached",
       color = "") +
  #transition_reveal(along = date, keep_last = T) +
  theme.maptheeme
```


## Cumulative number of cases registered outside of China

```{r}


coronavirus_ev <- 
coronavirus %>% 
  rename(Country = Country.Region) %>% 
  select(-Province.State, -Lat, -Long) %>% 
  group_by(Country, date, type) %>% 
  summarise(total = sum(cases))

coronavirus_ev %>% head() %>% kable()
```


```{r fig.height=11,fig.width=9, fig.align = "center", out.width="100%"}

library(ggflags)      # A flag geom for ggplot2
library(countrycode)  # Convert Country Names and Country Codes

coronavirus_ev <- 
coronavirus_ev %>% filter(Country != "Mainland China" ) 

top_three <- 
coronavirus_ev %>%  
  ungroup() %>%  
  filter(date == max(date)) %>%
  group_by(type) %>% 
  arrange(type, desc(total)) %>% 
  top_n(n = 3, wt = total) %>% 
  mutate(Country = ifelse(Country == "UK", "United Kingdom",
                          no = ifelse(Country == "US", "United States", no = Country)),
         rank = rank(desc(total)))


countrynames <- 
  codelist %>% 
  select(country.name.en, ecb) %>% 
  rename(Country = country.name.en)

top_three <- 
merge(top_three, countrynames, by = "Country", all.x = TRUE)


ggplot(data = coronavirus_ev, aes(x = date, y = total, group = Country)) +
  geom_line(aes(color = Country)) +
  geom_segment(data = top_three,
               aes(x = date, y = total, group = Country, xend = date + 4, yend = total),
               linetype = 2, size = 0.5) +
  geom_label(data = top_three,
             aes(x = date + 2, y = total, group = Country, label = total)) +
  geom_flag(data = top_three, 
            aes(x= date + 4, y = total, country = tolower(ecb)), size= 10) +
  facet_wrap(facets = .~type, ncol = 1, scales = "free_y") +
  guides(color=guide_legend(ncol=7,title.position = "bottom", title.hjust = 0.5)) +
  theme(legend.position = "bottom")
```


## Cases Distribution by Type 

```{r fig.height=11,fig.width=9, fig.align = "center", out.width="100%"}
fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}

coronavirus %>% 
  rename(Country = Country.Region) %>% 
  filter(Country != "Others") %>% 
  select(-Province.State, -Lat, -Long) %>% 
    mutate(Country = ifelse(Country == "UK", "United Kingdom",
                          no = ifelse(Country == "US", "United States", no = Country))) %>% 
  group_by(Country, type) %>% 
  summarise(total = sum(cases)) %>% 
  top_n(25, total) %>% 
ggplot(aes(x = total, y = reorder(x = Country,X =  total, FUN = max), fill = type)) +
  geom_bar(stat = "identity") +
    facet_grid(.~type, scales = "free_x") +
  scale_x_log10(labels = scales::label_number_si())
  

```


But as seen all the points are overlapped, because the date is not being considered. Now we are going to include this variable (time) as a _Dynamic_ plot, this using `library(gganimate)`. A very good guide explaining the features of this library is found [here](https://goodekat.github.io/presentations/2019-isugg-gganimate-spooky/slides.html).


## Total cases by country until today facet (bar plot)

Talk about geom_flag

Max bar plot use dot line

