---
title: "Data visualization with R coronavirus and ggplot2 Packages"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Coronavirus

As we know coronavirus is a huge problem currently and I found an oportunnity to enhance my programming skills in *R* with a simple dataset. I recently noticed Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) has released a package at CRAN which provides a daily summary of the **Coronavirus (COVID-19)** cases by state/province. The package is available at <https://cran.r-project.org/web/packages/coronavirus/index.html>. Below, I'll show some simple analysis with this dataset with the help of *ggplot2* in *R*, my aim for next posts is to show you how create dynamics plots with *gganimate* package.

The main gist of the process is to show how coronavirus is spread in the World nowadays, and how the cases in each country are increasing. To get started, you’ll need to download coronavirus from Github using the devtools package:

```{r, echo=T, message=F}
if (!require(devtools)) {
    install.packages("devtools")
}
if (!require(coronavirus)) {
  devtools::install_github(" RamiKrispin/coronavirus")
}
```

In addition to this package, I’ll also use `dplyr` and `magrittr` for data manipulation, `ggplot2` for plotting, and `maps` for geographical data. 

```{r , echo=T, message=F}
library(coronavirus)# The 2019 Novel Coronavirus COVID-19 (2019-nCoV) Dataset
library(dplyr)      # A Grammar of Data Manipulation
library(magrittr)   # A Forward-Pipe Operator for R
library(maps)       # Draw Geographical Maps
library(ggplot2)    # Create Elegant Data Visualisations Using the Grammar of Graphics
```

## COVID-19 spread evolution around the Globe

When loading the `library(coronavirus)` we get the dataset which contains the daily summary of Coronavirus cases (confirmed, death, and recovered), by state/province for some countries as seen below:

```{r, layout="l-body-outset"}
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
kable(head(coronavirus))
```

Filtering by "confirmed " cases and renaming some columns we get:

```{r}
coronavirus_data <- 
coronavirus %>% 
  filter(type == "confirmed") %>% 
  rename(Country = Country.Region)

coronavirus_data %>% head() %>% kable()
```

The next step is to plot each point on the globe:

```{r}
map.world <- map_data("world")

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group),
               fill="lightgray", colour = "gray50", size = 0.2) +
  geom_point(data = coronavirus_data, aes(x = Long, y = Lat, size = cases), 
             shape = 21, stroke =0.1, fill = "#ccff00") +
  labs(x="", y="", title = "COVID-19", subtitle = "Coronavirus evolution around the World",
       size = "Confirmed \ncases") +
  coord_fixed() +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))
```

But as seen all the points are overlapped, because the date is not being considered. Now we are going to include this variable (time) as a _Dynamic_ plot, this using `library(gganimate)`. A very good guide explaining the features of this library is found [here](https://goodekat.github.io/presentations/2019-isugg-gganimate-spooky/slides.html).

## Total cases until today World (facet by case)

## Total cases by country until today facet (bar plot)

Talk about geom_flag

Max bar plot use dot line


```{r, echo=FALSE}
ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group),
               fill="lightgray", colour = "gray50", size = 0.2) +
  geom_point(data = coronavirus, aes(x = Long, y = Lat, size = cases), alpha = 0.7,
             shape = 21, stroke =0.1, fill = "#ccff00") +
  coord_fixed() +
  labs(x="", y="", title = "COVID-19", subtitle = "Coronavirus evolution around the World",
       size = "Confirmed \ncases") +
    scale_size(range = c(5,7)) +
  #transition_time(date) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))
```

```{r}
Wuhan <- 
data.frame(Province.State = "Hubei",
           Lat = 30.97564,
           Long = 112.2707)

migr_corona <- 
coronavirus %>% filter(Province.State != "Hubei") %>% 
  select(Country.Region,date, Lat, Long) %>% distinct()

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group),
               fill="lightgray", colour = "gray50", size = 0.2) + 
  geom_curve(data=migr_corona,
             aes(x=Wuhan$Long, y=Wuhan$Lat, xend=Long, yend=Lat, color = Country.Region, group =date),
             # col="#00008b",
             size=.5,
             curvature=0.2) +
  geom_point(data=Wuhan,
             aes(x=Long, y=Lat), 
             fill="red", shape = 21, stroke=0.2,
             color ="white", size=1.5) +
  geom_point(data=migr_corona,
             aes(x=Long, y=Lat), 
             fill="gray40", shape = 21, stroke=0.2,
             color ="gray40",size=1.5) +
  coord_fixed() +
  guides(color=guide_legend(nrow = 3)) +
  labs(x="", y="", title = "COVID-19", subtitle = "Countries which have confirmed coronavirus cases",
       color = "") +
  #transition_reveal(along = date, keep_last = T) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"), 
        legend.position = "bottom")
```
```{r}
# ggplot(data = coronavirus_data, aes(group = date)) +
#   geom_polygon(data = map.world, aes(x = long, y = lat, group = group),
#                fill="lightgray", colour = "gray50", size = 0.2) +
#   geom_point(aes(x = Long, y = Lat, size = cases), alpha = 0.7,
#              shape = 21, stroke =0.1, fill = "#ccff00") +
#   geom_text(aes(x = Long, y = Lat, label = cases), hjust=0.5, vjust=0.2, size = 1.5) +
#   coord_fixed() +
#   transition_states(states = date,
#                     transition_length = 3,
#                     state_length = 3) +
#   enter_fade() +
#   exit_fade() +
#   labs(x="", y="", title = "COVID-19", subtitle = "Coronavirus evolution around the World",
#        caption = 'Year: ',
#        size = "Confirmed \ncases") +
#   theme_bw() +
#   theme(plot.title = element_text(face = "bold"))
```

